2.3 Prob Model for PCR Positivity Estimation

### 2024-04-11 2_scripts/2.3 Prob Model for PCR pos.qmd
Update Roche~Abbott models

# create a subset of APL
APLs<-RFD4682_e %>%
  select(order_ID, AGE_AT_COLLECTION, GENDER, N, Col_Month, PAT3)

# unify colnames
colnames(APLs)<-c("order_ID","age","gender","abbott_n","samplemonth","fsa")
APLs$samplemonth<-as.yearmon(APLs$samplemonth)

@@

#### 3.1 The linear (log) model
```{r}
# The linear model using log_e transformed abbott_n and roche_n

@@

#### 3.2 Apply the model lm01 to the CBS data
```{r}
log_abbott_n<-predict(lm01, newdata = CBSs, "response")
summary(log_abbott_n)

@@

#### 3.3 plot: sero-prevalence, using predicted Abbott positivity
```{r}
tb_abb<-CBSs %>%
  group_by(samplemonth, region) %>%
  summarise(n = n() , 
            pos_abbott= sum(pos_abbott, na.rm = T), 
            pos_roche= sum(pos_roche, na.rm = T)
            )
tb_abb$prop.abbott<-tb_abb$pos_abbott/tb_abb$n
tb_abb$prop.roche<-tb_abb$pos_roche/tb_abb$n

@@

@@

# using our training data:
p04<-dta_all %>%
  .[!is.na(.$log_abbott_n) & !is.na(.$log_roche_n),] %>%
  ggplot(aes(x=log_roche_n, y=log_abbott_n, 
             study, colour=study)) +
  geom_point(aes(colour=as.factor(study))) +
  geom_vline(xintercept = 0, linewidth=1.0, linetype="dotted", col="red") +
  geom_hline(yintercept = log(1.4), linewidth=1.0, linetype="dotted", col="red") +
  xlab("Roche Anti-N S/Co ratio (log scale)") + 
  ylab("Abbott Anti-N S/Co ratio (log scale)") +
  labs(color='') +
  scale_colour_manual(labels = c("Jamaica (-)", "Osaka, JP (-)",
                                 "Osaka, JP (+)","UK (+)"), 
                      values = c("red", "skyblue", "darkgreen", "purple")) +
  theme_bw() + 
  theme(legend.position = "bottom")

@@

p05<-dta_CBS %>%
  .[!is.na(.$log_abbott_n) & !is.na(.$log_roche_n),] %>%
  ggplot(aes(x=log_roche_n, y=log_abbott_n, 
             study, colour=study)) +
  geom_point(aes(colour=as.factor(study))) +
  geom_vline(xintercept = 0, linewidth=1.0, linetype="dotted", col="red") +
  geom_hline(yintercept = log(1.4), linewidth=1.0, linetype="dotted", col="red") +
  xlab("Roche Anti-N S/Co ratio (log scale)") + 
  ylab("Abbott Anti-N S/Co ratio (log scale)") +
  labs(color='') +
  scale_colour_manual(labels = c("CBS","Jamaica (-)", "Osaka, JP (-)",
                                 "Osaka, JP (+)","UK (+)"), 
                      values = c("lightblue","red", "blue", "darkgreen", "purple")) +
  theme_bw() + 
  theme(legend.position = "bottom")

>>>>

# create a subset of APL
APLs<-RFD4682_e %>%
  select(order_ID, AGE_AT_COLLECTION, GENDER, N, `N-IgG_INTERP`, Col_Month, PAT3)

# unify colnames
colnames(APLs)<-c("order_ID","age","gender","abbott_n", "pos_abbott","samplemonth","fsa")
APLs$samplemonth<-as.yearmon(APLs$samplemonth)

@@

#### 3.1.1 The linear (log) model: Abbott ~ Roche
```{r}
# The linear model using log_e transformed abbott_n and roche_n

@@

#### 3.1.2  The linear (log) model: Roche ~ Abbott
```{r}
lm01r<-glm(log_roche_n ~ log_abbott_n, 
          family = "gaussian",
          data = dta_all, weights = sample_wt)
summary(lm01r) 
```

#### 3.2.1 Apply the model lm01 to the CBS data
```{r}
log_abbott_n<-predict(lm01, newdata = CBSs, "response")
summary(log_abbott_n)

@@

#### 3.2.2 Apply model lm01r to APL data
```{r}
APLs$log_roche_n<-predict(lm01r, newdata = APLs, "response")

APLs$roche_n_hat <- exp(APLs$log_roche_n)
summary(APLs$roche_n_hat)

# Apply Roche COI=1.0
APLs$pos_roche<-if_else(APLs$roche_n_hat<=1, 0, 1, missing = NA)
# recode Abbott pos as num
APLs$pos_abbott[APLs$pos_abbott=="Negative"]<-0
APLs$pos_abbott[APLs$pos_abbott=="Positive"]<-1
APLs$pos_abbott<-as.numeric(APLs$pos_abbott)
```

#### 3.3.1 plot: sero-prevalence, using predicted Abbott positivity
```{r}
tb_abb<-CBSs %>%
  group_by(samplemonth, region) %>%
  summarise(n = n() , 
            pos_abbott= sum(pos_abbott, na.rm = T), 
            pos_roche= sum(pos_roche, na.rm = T)
            )
tb_abb$prop.abbott<-tb_abb$pos_abbott/tb_abb$n
tb_abb$prop.roche<-tb_abb$pos_roche/tb_abb$n

@@

#### 3.3.2 plot: sero-prevalence, using predicted Roche pos
```{r}
tb_ro<-APLs %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            pos_abbott= sum(pos_abbott, na.rm = T), 
            pos_roche= sum(pos_roche, na.rm = T)
            )
tb_ro$prop.abbott<-tb_ro$pos_abbott/tb_ro$n
tb_ro$prop.roche<-tb_ro$pos_roche/tb_ro$n

# clean NAs
tb_ro<-tb_ro[!is.na(tb_ro$samplemonth) & 
             !tb_ro$pos_abbott==0, ]

tb_ro_melt<-reshape2::melt(select(tb_ro, -c(2:4)), 
                        id.vars = c("samplemonth"),
                        variable.name = "assay",
                        value.name = "prop")

p03r<-tb_ro_melt %>%
  ggplot(aes(x=samplemonth, y=prop, group=assay, colour=assay)) +
  geom_point(aes(colour=assay)) +
  geom_line(aes(colour=assay))  +
  theme_bw() 
p03r
```

@@

# using our training data:
p04<-dta_all %>%
  .[!is.na(.$log_abbott_n) & !is.na(.$log_roche_n),] %>%
  ggplot(aes(x=log_roche_n, y=log_abbott_n, 
             study, colour=study)) +
  geom_point(aes(colour=as.factor(study)), alpha=0.5) +
  geom_vline(xintercept = 0, linewidth=1.0, linetype="dotted", col="red") +
  geom_hline(yintercept = log(1.4), linewidth=1.0, linetype="dotted", col="red") +
  xlab("Roche Anti-N S/Co ratio (log scale)") + 
  ylab("Abbott Anti-N S/Co ratio (log scale)") +
  labs(color='') +
  scale_colour_manual(labels = c("Jamaica (-)", "Osaka, JP (-)",
                                 "Osaka, JP (+)","UK (+)"), 
                      values = c("red", "blue", "darkgreen", "purple")) +
  theme_bw() + 
  theme(legend.position = "bottom")

@@

p05<-dta_CBS %>%
  .[!is.na(.$log_abbott_n) & !is.na(.$log_roche_n),] %>%
  ggplot(aes(x=log_roche_n, y=log_abbott_n, 
             study, colour=study)) +
  geom_point(aes(colour=as.factor(study)), alpha=0.40) +
  geom_vline(xintercept = 0, linewidth=1.0, linetype="dotted", col="red") +
  geom_hline(yintercept = log(1.4), linewidth=1.0, linetype="dotted", col="red") +
  xlab("Roche Anti-N S/Co ratio (log scale)") + 
  ylab("Abbott Anti-N S/Co ratio (log scale)") +
  labs(color='') +
  scale_colour_manual(labels = c("CBS","Jamaica (-)", "Osaka, JP (-)",
                                 "Osaka, JP (+)","UK (+)"), 
                      values = c("lightblue","red", "blue", "darkgreen", "purple")) +
  theme_bw() + 
  theme(legend.position = "bottom")
